Numerical simulation and entropy dissipative cure of the carbuncle instability for the shallow water circular hydraulic jump

### 2020-11-04 numerics_paper/version2/numerical_chj.tex
tmp update

%\subsection{Roe's Riemann solver}
\begin{align*}
  F_{i+1/2}=\frac{\bff(u_{i+1})+\bff(u_i)}{2}+\frac{1}{2}|\hat A_{i+1/2}|(u_{i+1}-u_i)
\end{align*}

@@

To understand the effect of choosing $s_{i+1/2}^p=\hat\lambda_{i+1/2}^p$ as the speed of $\W_{i+1/2}^p$, 
let us assume that no limiters are applied; i.e., $\tilde\W_{i\pm 1/2}^p\equiv \W_{i\pm 1/2}^p$.
By doing this, equation XX can be written as 
\begin{align*}
  u_i^{n+1}=u_i^n
  &-\frac{\Delta t}{2 \Delta x}
  \sum_p
  \left[\left(\hat\lambda_{i+1/2}^p -\frac{\Delta t}{\Delta x}|\hat\lambda_{i+1/2}^p|^2 \right)\W_{i+1/2}^p
  +
  \left(\hat\lambda_{i-1/2}^p +\frac{\Delta t}{\Delta x}|\hat\lambda_{i-1/2}^p|^2 \right)\W_{i-1/2}^p\right] \\
  &-\frac{\Delta t}{2\Delta x}
  \sum_p 
  \left[-\left(\lambda_{i+1/2}^{\max}-|\hat\lambda_{i+1/2}^p|\right)R_{i+1/2}\W_{i+1/2}^p
    +\left(\lambda_{i-1/2}^{\max}-|\hat\lambda_{i-1/2}^p|\right)R_{i-1/2}\W_{i-1/2}^p\right],
\end{align*}
where the first row corresponds to the standard Lax-Wendroff method written in terms of 
fluctuation-like quantities. 
Since $\lambda_{i\pm1/2}^{\max}\geq |\hat\lambda_{i\pm1/2}^p|$, 
the second row adds extra stabilization.
Note that if $R_{i\pm 1/2}=0$, we recover the standard second-order Lax-Wendroff method based on 
Roe's Riemann solver. 
In the numerical experiments in \S XX, we obtain second-order experimental 
order of convergence (EOC) for problems with smooth solution, which is expected since 
$R_{i\pm 1/2}\approx 0$ in smooth regions.
Also note that if $R_{i\pm 1/2}\approx 1$, which we expect to happen near steep gradients, 
we do not recover the Lax-Wendroff method based on Rusanov's Riemann solver. 
Instead, we obtain the first-order method XX with a flux correction. Assuming 
the CFL-like condition $\Delta t/\Delta x|\hat\lambda_{i+1/2}^p|-1\leq 0$, the flux correction is anti-diffusive, 
which slightly improves the accuracy of the first-order Rusanov's method. 
In our numerical experiments 
with smooth solutions, we obtain first-order EOC if we set $R_{i\pm1/2}=1$, $\forall i$. 
%
Therefore, the accuracy properties of method XX depend upon the behavior of $R_{i\pm 1/2}$. 
In smooth regions, where $R_{i\pm 1/2}\approx 0$, we recover second-order convergence;
near discontinuities, where $R_{i\pm 1/2}\approx 1$, we add extra dissipation that decreases the convergence 
to first-order. It is this extra dissipation that reduces the occurance of carbuncle instabilities 
in our numerical experiments. In smooth regions, the accuracy of the underlying Lax-Wendroff-LeVeque method 
based on Roe's Riemann solver is preserved.

>>>>

\subsection*{Roe's Riemann solver}
\begin{align*}
  F_{i+1/2}=\frac{\bff(u_{i+1})+\bff(u_i)}{2}+\frac{1}{2}|\hat A_{i+1/2}|(u_{i+1}-u_i)
\end{align*}

@@

To understand the effect of choosing $s_{i+1/2}^p=\hat\lambda_{i+1/2}^p$ as the speed of $\W_{i+1/2}^p$, 
let us assume that no limiters are applied; i.e., $\tilde\W_{i\pm 1/2}^p\equiv \W_{i\pm 1/2}^p$.
By doing this, equation XX can be written as 
\begin{align*}
  u_i^{n+1}=u_i^n
  &-\frac{\Delta t}{2 \Delta x}
  \sum_p
  \left[\left(\hat\lambda_{i+1/2}^p -\frac{\Delta t}{\Delta x}|\hat\lambda_{i+1/2}^p|^2 \right)\W_{i+1/2}^p
  +
  \left(\hat\lambda_{i-1/2}^p +\frac{\Delta t}{\Delta x}|\hat\lambda_{i-1/2}^p|^2 \right)\W_{i-1/2}^p\right] \\
  &-\frac{\Delta t}{2\Delta x}
  \sum_p 
  \left[-R_{i+1/2}\left(\lambda_{i+1/2}^{\max}-|\hat\lambda_{i+1/2}^p|\right)\W_{i+1/2}^p
    +R_{i-1/2}\left(\lambda_{i-1/2}^{\max}-|\hat\lambda_{i-1/2}^p|\right)\W_{i-1/2}^p\right],
\end{align*}
where the first row corresponds to the standard Lax-Wendroff method written in terms of 
fluctuation-like quantities. 
Since $\lambda_{i\pm1/2}^{\max}\geq |\hat\lambda_{i\pm1/2}^p|$, 
the second row adds extra stabilization.
Note that if $R_{i\pm 1/2}=0$, we recover the standard second-order Lax-Wendroff method based on 
Roe's Riemann solver. 
In the numerical experiments in \S XX, we obtain second-order experimental 
order of convergence (EOC) for problems with smooth solution, which is expected since 
$R_{i\pm 1/2}\approx 0$ in smooth regions.
Also note that if $R_{i\pm 1/2}\approx 1$, which we expect to happen near steep gradients, 
we do not recover the Lax-Wendroff method based on Rusanov's Riemann solver. 
Instead, we obtain the first-order method XX with a flux correction. Assuming 
the CFL-like condition $\Delta t/\Delta x|\hat\lambda_{i\pm1/2}^p|-1\leq 0$, the flux correction is anti-diffusive, 
which slightly improves the accuracy of the first-order Rusanov's method. 
In our numerical experiments 
with smooth solutions, we obtain first-order EOC if we set $R_{i\pm1/2}=1$, $\forall i$. 
%
Therefore, the accuracy properties of method XX depend upon the behavior of $R_{i\pm 1/2}$. 
In smooth regions, where $R_{i\pm 1/2}\approx 0$, we recover second-order convergence;
near discontinuities, where $R_{i\pm 1/2}\approx 1$, we add extra dissipation that decreases the convergence 
to first-order. It is this extra dissipation that reduces the occurance of carbuncle instabilities 
in our numerical experiments.
%In smooth regions, the accuracy of the underlying Lax-Wendroff-LeVeque method 
%based on Roe's Riemann solver is preserved.
